LDpred2: better, faster, stronger

### 2020-04-26 paper/paper-ldpred2.tex
save

\abstract{

}

@@

LDpred is a popular and powerful method for deriving polygenic scores based on summary statistics and a Linkage Disequilibrium (LD) matrix only \cite[]{vilhjalmsson2015modeling}.
It assumes there is a proportion $p$ of variants that are causal.
However, LDpred has several limitations that may result in limited predictive performance.
The non-infinitesimal version of LDpred, a Gibbs sampler, is particularly sensitive to model misspecification when applied to summary statistics with large sample sizes.
It is also unstable in long-range LD regions such as the human leukocyte antigen (HLA) region of chromosome 6. 
This issue has lead to the removal of such regions from analyses  \cite[]{marquez2018modeling,lloyd2019improved}, which is unfortunate since this region of the genome contains many known disease-associated variants, particularly with autoimmune diseases and psychiatric disorders \cite[]{mokhtari2016major,matzaraki2017mhc}.
%LDpred is also considered not particularly fast, nor memory-efficient, nor it provides truly sparse models as expected by its model specification.

@@

Here we present LDpred2, a new version of LDpred \cite[]{vilhjalmsson2015modeling}.
LDpred2 has 4 options: 1) LDpred2-inf, an infinitesimal model similar to the one in LDpred1; 2) LDpred2-grid (or simply LDpred2) that is the main LDpred model, where a grid of values for hyper-parameters $p$ (the proportion of causal variants) and $h^2$ (the SNP heritability) are tuned using a validation test; 3) LDpred2-sparse, which is similar to LDpred2-grid but where effects can be exactly 0, offering a truly sparse version of LDpred; 4) LDpred2-auto, which automatically estimate $p$ and $h^2$ and therefore is free of hyper-parameters to optimize.
As a recall, LDpred v1 has two options: LDpred1-grid where only $p$ is optimized, and LDpred1-inf.

We compare the two versions of LDpred using six simulation scenarios to understand in which cases it is important to use the new version. We also compare these two versions of LDpred using eight binary phenotypes of major interest and for which there are published external summary statistics available and a substantial number of cases in the UK Biobank data. Area Under the ROC Curve (AUC) values are reported.

@@

Figure \ref{fig:AUC-simu} presents the simulation results comparing LDpred1 (v1.0.0 as implemented by \cite{vilhjalmsson2015modeling}) with the new LDpred2 (as implemented in R package bigsnpr). 
six simulation scenarios are used, each repeated 10 times.
In the first four simulation scenarios, a heritability $h^2$ of 40\% and a prevalence of 15\% are used. We simulate 300, 3000, 30,000 or 300,000 causal variants anywhere on the genome. 
In these scenarios, infinitesimal models perform similarly. 
When testing a grid of hyper-parameters, LDpred2 performs substantially better than LDpred1 in the cases where the number of causal variants is smaller, i.e.\ in the case of 300 or 3000 causal variants ($p$=2.5e-4 and 2.5e-3). 
For example, in simulations with 300 causal variants, a mean AUC of 73.5\% is obtained with LDpred1 while a value of 81.9\% is obtained with LDpred2.
In these scenarios, all 3 non-infinitesimal models of LDpred2 perform equally well.
In HLA scenarios, i.e.\ when causal variants are sampled in a region with variants in high LD, LDpred1 performs poorly. In these two HLA scenarios, all LDpred2 models perform well, although LDpred2-auto performs slightly worse than other models.

@@

Figure \ref{fig:AUC-real} presents the results of real data applications comparing LDpred1 (v1.0.0 as implemented by \cite{vilhjalmsson2015modeling}) with the new LDpred2 (as implemented in R package bigsnpr). 
eight binary phenotypes are used, summarized in table \ref{tab:sumstats}.
For T1D, T2D, BRCA, PRCA and MDD, all LDpred2 models perform much better than LDpred1. For example, for BRCA, AUC improves from 58.5\% with LDpred1 to 65\% with LDpred2, and from 55\% to 75\% for T1D. 
For Asthma, CAD and RA, predictive performances of LDpred1 and LDpred2 are similar.

As in simulations, the sparse version of LDpred2 performs equally well as the original version, with the exception of PRCA, for which it performs poorly.
For CAD, PRCA and TD1, LDpred2-auto perform much worse than LDpred2, otherwise it performs similarly.

@@

In LDpred2, we also test more values for $p$ (17 instead of 7 by default) and for $h^2$ (3 instead of 1).
When testing the grid of hyper-parameters of $p$ and $h^2$, we also allow for testing an option to enable sparse models in LDpred2 (see after).
Overall, we test a grid of 102 different values in LDpred2 instead of 7 in LDpred1, we use 5 times as many iterations in the Gibbs sampler, we use a larger window size for computing correlation, yet LDpred2 is still as fast as LDpred1.
LDpred2 is fast because we provide an efficient parallel implementation in C++. The parallelization is performed on the grid of hyper-parameters, but can also be performed chromosome-wise, as each chromosome is processed independently in LDpred2 (see Methods).
As an example, it takes less than two hours to run LDpred2-inf, LDpred2-grid (108 hyper-parameter values) and LDpred2-auto for a chromosome with 100K variants and using 8 cores.
Pre-computing the LD matrix for a chromosome with 100K variants takes 11 minutes using 16 cores.

LDpred2 also comes with two new models.
When the sparse option is enabled, it provides models that truly encourage sparsity, as compared to LDpred1 which outputs very small non-zero effect sizes \cite[]{cecile2019polygenic}.
It can provide sparsity in effect sizes as much as 98\% for small $p$, while keeping predictive performance as good as non-sparse models, as opposed to if very small effects were simply discarded \cite[]{bolli2019software}.
This sparse model performs equally well as the non-sparse model, therefore we encourage users to test it in the grid and to choose it if it performs equally well as the non-sparse model.
As for LDpred2-auto, which automatically estimates values for hyper-parameters $p$ and $h^2$, it performs equally well as other LDpred2 models in simulations, but does not perform well for half of the real traits analyzed here.
More investigations need to be performed to understand the poor results of LDpred2-auto for these particular real traits.

\begin{figure}[h]
	\centerline{\includegraphics[width=0.85\textwidth]{QC-CAD}}
	\caption{Results of the real data applications using published external summary statistics.
		Bars present AUC values on the test set of UKBB (mean and 95\% CI from $10^4$ bootstrap samples).
		%See corresponding values in table S2.
	}
	\label{fig:QC-CAD}
\end{figure}

[TALK ABOUT QUALITY CONTROL ON AF?]

[TALK ABOUT DIFFERENT SETS OF VARIANTS?]

[NEED TO UNDERSTAND WHY AUTO DOES NOT WORK FOR SOME TRAITS -> CAD BECAUSE DIFFERENT N // T1D PROBABLY BECAUSE OF SMALL SAMPLE SIZE // ASTHMA? AND PRCA?]

[TALK ABOUT MODEL FOR BINARY DATA]

@@

We use the UK Biobank data for both real data analyses and simulations \cite[]{bycroft2018uk}. We restrict individuals to White-British individuals used in the PC computation of the UK Biobank (i.e.\ unrelated and quality-controlled). We restrict variants to HapMap3 variants (considered to be well imputed).
This results in 337,475 individuals and 1,194,574 variants.
We use 10,000 individuals as a validation set for choosing optimal hyper-parameters and for computing correlations between variants (LD matrix). 
We use 300,000 other individuals for running logistic GWAS to create summary statistics.
We use the remaining 27,475 individuals as test set for evaluating models.

@@

%\begin{minipage}{.7\textwidth}
\begin{algorithm}[H]
\caption{LDpred, with hyper-parameters $p$ and $h^2$}\label{algo:ldpred}
\begin{algorithmic}[1]
	\State $\boldsymbol{\hat{\beta}} \gets \dfrac{\boldsymbol{\hat{\gamma}}}{\boldsymbol{\text{se}(\hat{\gamma})} \cdot \sqrt{\boldsymbol{N}}}$ \Comment{See previous section}
	\State $\boldsymbol{\Omega} \gets \boldsymbol{0}$
	\For {$k=1,\ldots,N_\text{burn-in} + N_\text{iter}$} \Comment{Gibbs iterations}
	\For {each variant $j$} \Comment{All variants}
	\State Compute $\tilde{\beta}_j$ according to \eqref{eq:beta_res}
	\State Compute $\bar{p}_j$ according to \eqref{eq:postp}
	\State Sample $\beta_j$ according to \eqref{eg:random_beta}
	\State Compute $\omega_j$ according to \eqref{eg:post_beta}
	\EndFor
	\State Compute $\boldsymbol{\beta}^T \boldsymbol{K} \boldsymbol{\beta}$ (estimate of $h^2$) to check divergence \Comment{Return 0s if divergence}
	\If {$k > N_\text{burn-in}$} 
	\State Increment $\boldsymbol{\Omega}$ with $\boldsymbol{\omega}$ 
	\EndIf
	\EndFor
	\State $\boldsymbol{\Omega} \gets \boldsymbol{\Omega} /  N_\text{iter}$ \Comment{Average of all $\boldsymbol{\omega}$ after burn-in}
	\State Return $\boldsymbol{\Omega} \cdot  \boldsymbol{\text{se}(\hat{\gamma})} \cdot \sqrt{\boldsymbol{N}}$ \Comment{See previous section}
\end{algorithmic} 
\end{algorithm}
%\end{minipage}

[CHANGE NOTATION OF SUMSTATS?]

@@

The first extension consists in estimating $p$ and $h^2$ in the model, as opposed to testing several values of $p$ and estimating $h^2$ using constrained LD score regression \cite[]{bulik2015ld}. This makes LDpred2-auto a method free of hyper-parameters which can therefore directly be applied to data without the need of a validation dataset to choose best-performing hyper-parameters.
To estimate $p$, we count the number of causal variants (i.e.\ $m_c = \sum_j(\beta_j \neq 0)$ in equation \ref{eg:random_beta}). 
We can assume that $m_c \sim \text{Binom}(M, p)$, so if we place a prior $p \sim \text{Beta}(1, 1) \equiv \mathcal{U}(0, 1)$, we can sample $p$ from the posterior $p \sim \text{Beta}(1 + m_c, 1 + M - m_c)$.
Due to complexity reasons, we could not derive a Bayesian estimator of $h^2$. Instead, we estimate $h^2 = \beta^T K \beta$, where $K$ is the correlation matrix.
These parameters $p$ and $h^2$ are updated after the inner loop in algorithm \ref{algo:ldpred}, then these new values are used in the next iteration of the outer loop.

The second extension, LDpred2-sparse, aims at providing sparse effect size estimates, i.e.\ that some resulting effects are exactly 0.
When the sparse solution is sought and when $\bar{p}_j < p$, we set $\beta_j$ and $\omega_j$ to 0 (lines 6-8 of algorithm \ref{algo:ldpred}).
Note that LDpred2-auto does not have a sparse option, but it is possible to run LDpred2-sparse with the $p$ and $h^2$ estimates from LDpred2-auto.

@@

LDpred has a window size parameter that needs to be set; for a given variant, correlations with other variants outside of this window are assumed to be 0.
The recommended value for this window (in number of variants) is to use the total number of variants divided by 3000, which corresponds to a window radius of around 2 Mb \cite[]{vilhjalmsson2015modeling}.
We have come to the conclusion that this window size is not large enough. Indeed, the human leukocyte antigen (HLA) region of chromosome 6 is 8 Mb long \cite[]{price2008long}. 
Using a window of 8Mb would be computationally and memory inefficient.
Instead, we propose to use genetic distances. Genome-wide, 1 Mb corresponds on average to 1 cM. Yet, the HLA region is only 3 cM long (vs.\ 8 Mb long).
Therefore, genetic distances enable to capture the same LD using a globally smaller window.
We provide function \texttt{snp\_asGeneticPos} in package bignspr to easily interpolate physical position to genetic positions.
We recommend to use genetic positions and to use a size parameter of 3 cM when computing the correlation between variants for LDpred2. 
Note that, in the code, we use \texttt{size = 3 / 1000} since parameter \texttt{size} is internally multiplied by 1000 in functions of package bigsnpr.
To prevent storing very small correlations, we also set to 0 all correlations with a p-value larger than \texttt{alpha = 0.9} (when testing that the correlation is different from 0).

@@

We recommend to run LDpred2 per chromosome.
Even if it is possible to run LDpred2 genome-wide, this approach has two limitations. First, it can be memory and computationally demanding to do so. For around one million (1M) variants, storing the 1M$\times$1M sparse correlation matrix takes more than 32 GB of memory. 
Doubling to 2M variants would require 128 GB of RAM to store the matrix.
Second, as noted in \cite{prive2019making}, it may be beneficial to assume that architecture of traits may be different for different chromosomes.
For example, chromosome 6 clearly encompasses a larger proportion of the heritability of autoimmune diseases compared to other chromosomes \cite[]{shi2016contrasting}.
Assuming the same model for genetic effects genome-wide could result in severe model misspecification, which would lead to suboptimal predictive performance.
Moreover, since the inverse of a block-diagonal matrix is formed from the inverse of each block, it should be safe to run LDpred2 for each chromosome and then combine the results.

@@

[TODO] 
All code used for this paper is available at \url{https://github.com/privefl/paper-ldpred2/tree/master/code}.
R package bigsnpr can be installed from either CRAN or GitHub (see \url{https://github.com/privefl/bigsnpr}).
A tutorial on the steps to run LDpred2 using some small example data is available at \url{https://privefl.github.io/bigsnpr/articles/LDpred2.html}.

>>>>

\abstract{	
In recent years the use of polygenic scores has become widespread.
LDpred is a popular and powerful method for deriving polygenic scores based on summary statistics and a Linkage Disequilibrium (LD) matrix only.
However, LDpred has several limitations that may result in limited predictive performance.
Here, we present LDpred2, a new version of LDpred that addresses these issues therefore providing higher predictive performance, while markedly improving its computational efficiency.
We provide this faster and more robust implementation of LDpred in R package bigsnpr, as well as two new options in LDpred2: ``sparse'' and ``auto''.
}

@@

LDpred is a popular and powerful method for deriving polygenic scores based on summary statistics and a Linkage Disequilibrium (LD) matrix only \cite[]{vilhjalmsson2015modeling}.
It assumes there is a proportion $p$ of variants that are causal.
However, LDpred has several limitations that may result in limited predictive performance.
The non-infinitesimal version of LDpred, a Gibbs sampler, is particularly sensitive to model misspecification when applied to summary statistics with large sample sizes.
It is also unstable in long-range LD regions such as the human leukocyte antigen (HLA) region of chromosome 6. 
This issue has led to the removal of such regions from analyses  \cite[]{marquez2018modeling,lloyd2019improved}, which is unfortunate since this region of the genome contains many known disease-associated variants, particularly with autoimmune diseases and psychiatric disorders \cite[]{mokhtari2016major,matzaraki2017mhc}.
%LDpred is also considered not particularly fast, nor memory-efficient, nor it provides truly sparse models as expected by its model specification.

@@

Here we present LDpred2, a new version of LDpred \cite[]{vilhjalmsson2015modeling}.
LDpred2 has 4 options: 1) LDpred2-inf, an infinitesimal model similar to the one in LDpred1; 2) LDpred2-grid (or simply LDpred2) that is the main LDpred model, where a grid of values for hyper-parameters $p$ (the proportion of causal variants) and $h^2$ (the SNP heritability) are tuned using a validation set; 3) LDpred2-sparse, which is similar to LDpred2-grid but where effects can be exactly 0, offering a version of LDpred that can provide sparse effects; 4) LDpred2-auto, which automatically estimate $p$ and $h^2$ and therefore is free of hyper-parameters to tune.
As a recall, LDpred v1 has two options: LDpred1-grid where only $p$ is optimized, and LDpred1-inf.

We compare the two versions of LDpred using six simulation scenarios to understand in which cases it is important to use the new version of LDpred. We also compare these two versions of LDpred using eight binary phenotypes of major interest and for which there are published external summary statistics available and a substantial number of cases in the UK Biobank data. Area Under the ROC Curve (AUC) values are reported.

@@

Figure \ref{fig:AUC-simu} presents the simulation results comparing LDpred1 (v1.0.0 as implemented by \cite{vilhjalmsson2015modeling}) with the new LDpred2 (as implemented in R package bigsnpr). 
Six simulation scenarios are used, each repeated 10 times.
In the first four simulation scenarios, a heritability $h^2$ of 40\% and a prevalence of 15\% are used. We simulate 300, 3000, 30,000 or 300,000 causal variants anywhere on the genome. 
In these scenarios, infinitesimal models perform similarly. 
When testing a grid of hyper-parameters, LDpred2 performs substantially better than LDpred1 in the cases where the number of causal variants is smaller, i.e.\ in the case of 300 or 3000 causal variants ($p$=2.5e-4 and 2.5e-3). 
For example, in simulations with 300 causal variants, a mean AUC of 73.5\% is obtained with LDpred1 while a value of 81.9\% is obtained with LDpred2.
In these scenarios, all 3 non-infinitesimal models of LDpred2 perform equally well.
In HLA scenarios, i.e.\ when causal variants are sampled in a region with variants in high LD, LDpred1 performs poorly. In these two HLA scenarios, all LDpred2 models perform well, although LDpred2-auto performs slightly worse than other models.

@@

Figure \ref{fig:AUC-real} presents the results of real data applications comparing LDpred1 (v1.0.0 as implemented by \cite{vilhjalmsson2015modeling}) with the new LDpred2 (as implemented in R package bigsnpr). 
Eight binary phenotypes are used, summarized in table \ref{tab:sumstats}.
For T1D, T2D, BRCA and PRCA, all main LDpred2 models perform much better than LDpred1. These improvements are also significant for MDD and CAD, albeit to a lesser extend. For example, for BRCA, AUC improves from 58.5\% with LDpred1 to 64.4\% with LDpred2, and from 54.9\% to 74.3\% for T1D. 
For Asthma and RA, predictive performances of LDpred1 and LDpred2 are similar.

As in simulations, the sparse version of LDpred2 performs equally well as the original version.
For PRCA and TD1, LDpred2-auto perform much worse than LDpred2, otherwise it performs similarly.

@@

In LDpred2, we also test more values for $p$ (17 instead of 7 by default) and for $h^2$ (3 instead of 1).
When testing the grid of hyper-parameters of $p$ and $h^2$, we also allow for testing an option to enable sparse models in LDpred2 (see after).
Overall, we test a grid of 102 different values in LDpred2 instead of 7 in LDpred1.
We also use 5 times as many iterations in the Gibbs sampler and a larger window size for computing correlations between variants, yet LDpred2 is still as fast as LDpred1.
LDpred2 is fast because we provide an efficient parallel implementation in C++. The parallelization is performed on the grid of hyper-parameters, but can also be performed chromosome-wise, as each chromosome is processed independently in LDpred2 (see Methods).
As an example, it takes around two hours in total to run LDpred2-inf, LDpred2-grid (108 hyper-parameter values parallelized over 8 cores) and LDpred2-auto for a chromosome with 100K variants.
Pre-computing the LD matrix for a chromosome with 100K variants takes 11 minutes using 16 cores.

LDpred2 also comes with two new models.
When the sparse option is enabled, it provides models that truly encourage sparsity, as compared to LDpred1 which outputs very small non-zero effect sizes \cite[]{cecile2019polygenic}.
It can provide sparsity in effect sizes as much as 98\% for small $p$, while keeping predictive performance as good as non-sparse models, as opposed to if very small effects were simply discarded \cite[]{bolli2019software}.
This sparse model performs equally well as the non-sparse model, therefore we encourage users to test it in the grid and to choose it if it performs better or equally well as the non-sparse model.
As for LDpred2-auto, which automatically estimates values for hyper-parameters $p$ and $h^2$, it performs equally well as other LDpred2 models in simulations, but does not perform well for some of the real traits analyzed here.
This is expected for T1D because it is mainly composed of large effects in the HLA region (see similar simulation results) and because summary statistics have a small sample size (5913 cases and 8828 controls only).
Yet, we do not know yet why LDpred2-auto performs poorly specifically for PRCA. More investigations need to be performed to understand these poor results, and to find if this issue can be fixed. Solutions that come to mind include for example running multiple chains with different initial values of $p$ and keeping the one with the best convergence properties.

%[TALK ABOUT QUALITY CONTROL ON AF?]

%[TALK ABOUT DIFFERENT SETS OF VARIANTS?]

%[TALK ABOUT MODEL FOR BINARY DATA]

@@

We use the UK Biobank data for both real data analyses and simulations \cite[]{bycroft2018uk}. We restrict individuals to White-British individuals used in the PC computation of the UK Biobank (i.e.\ unrelated and quality-controlled). We restrict variants to HapMap3 variants.
This results in 337,475 individuals and 1,194,574 variants.
We use 10,000 individuals as a validation set for choosing optimal hyper-parameters and for computing correlations between variants (LD matrix). 
We use 300,000 other individuals for running logistic GWAS to create summary statistics.
We use the remaining 27,475 individuals as test set for evaluating models.

@@

%\begin{minipage}{.7\textwidth}
\begin{algorithm}[H]
\caption{LDpred, with hyper-parameters $p$ and $h^2$, and summary statistics $\boldsymbol{\hat{\gamma}}$, $\boldsymbol{\text{se}(\hat{\gamma})}$ and $\boldsymbol{N}$}\label{algo:ldpred}
\begin{algorithmic}[1]
	\State $\boldsymbol{\hat{\beta}} \gets \dfrac{\boldsymbol{\hat{\gamma}}}{\boldsymbol{\text{se}(\hat{\gamma})} \cdot \sqrt{\boldsymbol{N}}}$ \Comment{See previous section}
	\State $\boldsymbol{\Omega} \gets \boldsymbol{0}$
	\For {$k=1,\ldots,N_\text{burn-in} + N_\text{iter}$} \Comment{Gibbs iterations}
	\For {each variant $j$} \Comment{All variants}
	\State Compute $\tilde{\beta}_j$ according to \eqref{eq:beta_res}
	\State Compute $\bar{p}_j$ according to \eqref{eq:postp}
	\State Sample $\beta_j$ according to \eqref{eg:random_beta}
	\State Compute $\omega_j$ according to \eqref{eg:post_beta}
	\EndFor
	\State Compute $\boldsymbol{\beta}^T \boldsymbol{K} \boldsymbol{\beta}$ (estimate of $h^2$) to check divergence \Comment{Return 0s if divergence}
	\If {$k > N_\text{burn-in}$} 
	\State Increment $\boldsymbol{\Omega}$ with $\boldsymbol{\omega}$ 
	\EndIf
	\EndFor
	\State $\boldsymbol{\Omega} \gets \boldsymbol{\Omega} /  N_\text{iter}$ \Comment{Average of all $\boldsymbol{\omega}$ after burn-in}
	\State Return $\boldsymbol{\Omega} \cdot  \boldsymbol{\text{se}(\hat{\gamma})} \cdot \sqrt{\boldsymbol{N}}$ \Comment{See previous section}
\end{algorithmic} 
\end{algorithm}
%\end{minipage}

%[CHANGE NOTATION OF SUMSTATS?]

@@

The first extension consists in estimating $p$ and $h^2$ in the model, as opposed to testing several values of $p$ and estimating $h^2$ using constrained LD score regression \cite[]{bulik2015ld}. This makes LDpred2-auto a method free of hyper-parameters which can therefore be applied directly to data without the need of a validation dataset to choose best-performing hyper-parameters.
To estimate $p$, we count the number of causal variants (i.e.\ $m_c = \sum_j(\beta_j \neq 0)$ in equation \ref{eg:random_beta}). 
We can assume that $m_c \sim \text{Binom}(M, p)$, so if we place a prior $p \sim \text{Beta}(1, 1) \equiv \mathcal{U}(0, 1)$, we can sample $p$ from the posterior $p \sim \text{Beta}(1 + m_c, 1 + M - m_c)$.
Due to complexity reasons, we could not derive a Bayesian estimator of $h^2$. Instead, we estimate $h^2 = \boldsymbol{\beta}^T \boldsymbol{K} \boldsymbol{\beta}$, where $\boldsymbol{K}$ is the correlation matrix.
These parameters $p$ and $h^2$ are updated after the inner loop in algorithm \ref{algo:ldpred}, then these new values are used in the next iteration of the outer loop.

The second extension, LDpred2-sparse, aims at providing sparse effect size estimates, i.e.\ some resulting effects are exactly 0.
When the sparse solution is sought and when $\bar{p}_j < p$, we set $\beta_j$ and $\omega_j$ to 0 (lines 6-8 of algorithm \ref{algo:ldpred}).
Note that LDpred2-auto does not have a sparse option, but it is possible to run LDpred2-sparse with the estimates of $p$ and $h^2$ from LDpred2-auto.

@@

LDpred has a window size parameter that needs to be set; for a given variant, correlations with other variants outside of this window are assumed to be 0.
The recommended value for this window (in number of variants) is to use the total number of variants divided by 3000, which corresponds to a window radius of around 2 Mb \cite[]{vilhjalmsson2015modeling}.
We have come to the conclusion that this window size is not large enough. Indeed, the human leukocyte antigen (HLA) region of chromosome 6 is 8 Mb long \cite[]{price2008long}. 
Using a window of 8Mb would be computationally and memory inefficient.
Instead, we propose to use genetic distances. Genome-wide, 1 Mb corresponds on average to 1 cM. Yet, the HLA region is only 3 cM long (vs.\ 8 Mb long).
Therefore, genetic distances enable to capture the same LD using a globally smaller window.
We provide function \texttt{snp\_asGeneticPos} in package bignspr to easily interpolate physical positions (in bp) to genetic positions (in cM).
We recommend to use genetic positions and to use a size parameter of 3 cM when computing the correlation between variants for LDpred2. 
Note that, in the code, we use \texttt{size = 3 / 1000} since parameter \texttt{size} is internally multiplied by 1000 in functions of package bigsnpr.
To prevent storing very small correlations, we also set to 0 all correlations with a p-value larger than \texttt{alpha = 0.9} (when testing that the correlation is different from 0).

@@

We recommend to run LDpred2 per chromosome.
Even if it is possible to run LDpred2 genome-wide, this approach has two limitations. First, it can be memory and computationally demanding to do so. For around one million (1M) variants, storing the 1M$\times$1M sparse correlation matrix takes more than 32 GB of memory. 
Doubling to 2M variants would require 128 GB of RAM to store the matrix.
Second, as noted in \cite{prive2019making}, it may be beneficial to assume that architecture of traits may be different for different chromosomes.
For example, chromosome 6 clearly encompasses a larger proportion of the heritability of autoimmune diseases compared to other chromosomes \cite[]{shi2016contrasting}.
Assuming the same model for genetic effects genome-wide could result in severe model misspecification, which would lead to suboptimal predictive performance.
Moreover, since the inverse of a block-diagonal matrix is formed from the inverse of each block, it should be safe to run LDpred2 for each chromosome and then combine the results. Combining results from all chromosomes requires using some appropriate scaling (see tutorial).

@@

%[TODO] 
%All code used for this paper is available at \url{https://github.com/privefl/paper-ldpred2/tree/master/code}.
The newest version of R package bigsnpr can be installed from GitHub (see \url{https://github.com/privefl/bigsnpr}).
A tutorial on the steps to run LDpred2 using some small example data is available at \url{https://privefl.github.io/bigsnpr/articles/LDpred2.html}.
